---
title: "R Notebook"
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_notebook: default
always_allow_html: yes
---

```{r setup, include=FALSE}

# This chunk shows/hides the code in your final report. When echo = TRUE, the code
# is shown in the report. When echo = FALSE, the code is hidden from the final report.
# We would like to see your code, so please leave the setting as is during the course.
# This chunk will not show up in your reports, so you can safely ignore its existence.

knitr::opts_chunk$set(echo = TRUE)

```


The following is your first chunk to start with. Remember, you can add chunks using the menu
above (Insert -> R) or using the keyboard shortcut Ctrl+Alt+I. A good practice is to use
different code chunks to answer different questions. You can delete this comment if you like.

Other useful keyboard shortcuts include Alt- for the assignment operator, and Ctrl+Shift+M
for the pipe operator. You can delete these reminders if you don't want them in your report.


```{r}
library("tidyverse")
library("tidymodels")
library("plotly")
library("skimr")
```
```{r}
library("gapminder")
dfGap <- gapminder
```

```{r}
#3a
skim(dfGap)
```

```{r}
#3b
dfGap %>%
  filter(year==2007) %>% 
  arrange(desc(lifeExp))
# The countries Japan, Hong Kong China, Iceland, Switzerland and Australia have life expectancy over 81.
```


```{r}
#3c
newdfGap<-
  dfGap %>%
  mutate(totalGDP=pop*gdpPercap) %>% 
  filter(year==2007) %>% 
  arrange(desc(totalGDP))
newdfGap


  dfGap %>%
  mutate(totalGDP=pop*gdpPercap) %>% 
  filter(year==2007) %>% 
  select(country,gdpPercap) %>% 
  arrange(desc(gdpPercap))


# The countries United States, China, Japan, India and Germany  are some of countries with highest totalGDP
# The countries Japan overlap with the countries in 3b
# The gdpPercapita for such countries is high but the population is low, therefore the totalGDP comes out to be less and therefore less overlap
```


```{r}
#3d
newdfGap %>% 
  group_by(continent) %>% 
  summarize(median_lifeExp=median(lifeExp), median_totalGDP = median (totalGDP)) %>% 
  ungroup() %>% 
  arrange(desc(median_lifeExp))

# The continent Oceania has the highest median life expectancy

```


```{r}
#4a
newdfGap %>% 
  ggplot(aes(x=totalGDP,y=lifeExp))+
  geom_point()
```


```{r}
newdfGap %>% 
  filter(totalGDP>1e+12) %>% 
  ggplot(aes(x=totalGDP,y=lifeExp))+
  geom_point()
```


```{r}

  newdfGap %>% 
  filter(totalGDP>1e+12) %>% 
  ggplot(aes(x=totalGDP,y=lifeExp))+
  geom_label(aes(label=country))+
  geom_point()
  
```


```{r}
library(ggrepel)
newdfGap %>% 
  filter(totalGDP>1e+12) %>% 
  ggplot(aes(x=totalGDP,y=lifeExp))+
  geom_label_repel(aes(label=country))+
  geom_point()

```


```{r}
newdfGap %>% 
  filter(totalGDP>1e+12) %>% 
  ggplot(aes(x=totalGDP,y=lifeExp,color=continent))+
  geom_label_repel(aes(label=country))+
  geom_point()


```


```{r}
#4b
newdfGap %>% 
  ggplot(aes(x=gdpPercap,y=lifeExp))+
  geom_line()+
  geom_smooth()
```


```{r}
#4c
boxPlotsForAll<-
  dfGap %>% 
  ggplot(aes(x=continent,y=lifeExp))+
  geom_boxplot()+
  ggtitle("Life expectancy across different continents")
boxPlotsForAll
```


```{r}
ggplotly(boxPlotsForAll)
```


```{r}
#4d
gdpOverTime<-dfGap %>% 
  group_by(continent,year) %>% 
  summarize(median_gdpPerCapita=median(gdpPercap)) %>% 
  ungroup() %>% 
  ggplot(aes(x=year,y=median_gdpPerCapita,color=continent))+
  geom_line()

gdpOverTime

```
```{r}
dfGap %>% 
  group_by(continent,year) %>% 
  summarize(mean_gdpPerCapita=mean(gdpPercap)) %>% 
  ungroup() %>% 
  ggplot(aes(x=year,y=mean_gdpPerCapita,color=continent))+
  geom_line()
```


```{r}
ggplotly(gdpOverTime)
```
